### 2017-08-03 paper/paper.tex
Updates.

\documentclass[pageno]{jpaper}

%replace XXX with the submission number you are given from the ASPLOS submission site.
\newcommand{\asplossubmissionnumber}{XXX}

\usepackage[normalem]{ulem}
\usepackage{parskip}

\begin{document}

\title{\Large \bf Cyclone: High Availability for Persistent Key Value Stores}

\date{}
\maketitle

\thispagestyle{empty}

\begin{abstract}
\vspace{0.1in}
Persistent key value stores are rapidly becoming an important component of many
distributed data serving solutions with innovations targeted at taking advantage
of growing flash speeds. Unfortunately their performance is hampered by the need
to maintain and replicate a write ahead log to guarantee availability in the
face of machine and storage failures. Cyclone is a replicated log plug-in for
key value stores that systematically addresses various sources of this
bottleneck. It uses a small amount of non-volatile memory directly addressable
by the CPU - such as in the form of NVDIMMs - to remove block oriented IO
devices such as SSDs from the critical path for appending to the log. It then
addresses network overheads using an implementation of the RAFT consensus
protocol that is designed around a userspace network stack to relieve the CPU of
the burden of data copies. Finally, it provides a way to efficiently map the
commutativity in key-value store APIs to the parallelism available in commodity
NICs via multiple transmit receive queues to scale replication performance on
multicore machines. Cyclone is able to replicate millions of small updates a
second using only commodity 10 gigabit ethernet adapters. As a practical
application, we use it to improve the performance and availability of RocksDB, a
popular persistent key value store by an order of magnitude when compared to its
own write ahead log without replication.
\end{abstract}

\section{Introduction}
Persistent key value stores are an increasingly important component of
datacenter scale storage services. Key value stores such as
Rocksdb~\cite{rocksdb}, LevelDB~\cite{leveldb} and FloDB~\cite{flodb} represent
large amounts of effort on both the engineering and research fronts. These key
value stores include sophisticated in-memory data structures built around Log
Structured Merge trees~\cite{lsmtree} and are heavily tuned to extract maximum
performance from flash-based solid state drives (SSDs).

These key value stores however tend to ignore an important component: the write
ahead log. A machine or storage failure leading to total loss or temporary
unavailability of data is unacceptable in services where high availabilty and
revenue are interconnected. Key value stores therefore usually incorporate
support for a write ahead log that if replicated and kept durable for every
appended update provides the necessary high availability. Unfortunately, the
write ahead log is a performance achilles heel for these systems - eclipsing
much of the work on improving the performance of the LSM component. To
illustrate the impact of the write ahead log, consider
Figure~\ref{fig:problem}. The line marked 'Rocksdb' shows the performance of
Rocksdb without the write ahead log. The performance when persisting the write
ahead log without replicating it is shown as the line marked 'rocksdb/WAL'. The
line marked 'rocksdb/3 way rep.' is for simply replicating the log three ways
without persisting it, using RAFT~\cite{raft} running over TCP/IP. Either
persisting every update to the log or replicating it using TCP/IP causes
performance to drop by an order of magnitude (note the log scale on the
x-axis). It is therefore no surprise that deployments of Rocksdb often turn off
the write ahead log~\cite{samza}, depending on upper layers to provide
availability by co-ordinating replicas. On the other side of the spectrum, key
value store research prototypes such as FloDB~\cite{flodb} turn off the write
ahead log to be able to showcase benefits of sophisticated extensions to LSM
datastructures.

\begin{figure}
\centering \includegraphics[scale=0.6]{results2/problem.pdf}
\caption{Rocksdb write ahead logging impact}
\label{fig:problem}
\end{figure}

Cyclone is a high speed strongly consistent replicated write ahead logging
service specialized for key value stores such as Rocksdb. Cyclone entirely
closes the performance gap due to both the storage and the network shown in
Figure~\ref{fig:problem}. We show that Cyclone achieves performance transparent
replication in that it can provide both a persistent write ahead log and
replication without compromising on RocksDB's performance.

The first key contribution of Cyclone is demonstrating the benefit of a small
amount of non-volatile memory when combined with the flash SSD based persistent
logs.  The primary cause of slowdown when turning on a write ahead log on a
single machine in Figure~\ref{fig:problem} is the need to do synchronous block
IO for every update \emph{regardless of the size of the update}. SSDs provide
the maximum throughput when writing large chunks (usually more than 4KB) of
data. This is due to the fixed cost of a round trip to the SSD through the IO
interface (NVMe in our case), regardless of the volume of data being
synchronously persisted. Cyclone avoids this by making use of directly attached
persistent memory on the server. This can be in the form of NVDIMMs (DIMMs with
an added ultracapacitor to dump state to a small amount of attached flash) or
newer forms of persistent memory~\cite{pmfs}. For current deployments, NVDIMMs
are significantly more expensive than standard DIMMs. This means that they will
likely be available in small quantities. We compensate for this by periodically
draining the NVDIMM log into a log placed on a standard SSD on the system.

The second key contribution of this work is in showing that the throughput
overheads in the network stack can be mitigated entirely in software. Existing
work on log replication using state machine replication protocols such as
Paxos~\cite{paxos} and RAFT~\cite{raft} has not - thus far - addressed high
performance replication in the local area network of a datacenter using
commodity networks.  Research focusing on addressing wide area network latencies
by reducing the number of network hops with new protocols such as Fast
Paxos~\cite{fast-paxos} or network switch modifications such as No
Paxos~\cite{nopaxos} do not address throughput bottlenecks due to network stack
related inefficiencies in the participating nodes of the quorum - a more urgent
problem with local area replication. On the other hand, work done to address
network related inefficiencies start with the assumption that the bottleneck to
good throughput on the network is the CPU and therefore one must either resort
to offload mechanisms such as RDMA~\cite{dare, farm, faast} or offload the
protocol entirely to an FPGA~\cite{consensus_box}. We demonstrate in this paper
that this is not the case. Rather, once persistence is provided by directly
attached non-volatile memory eliminating block IO from the persistence step in
replication protocols, the problem becomes equivalent to multicasting the same
log entry from the leader to all followers. We design Cyclone's network stack
around well known principles in the software packet switching community. In
particular, a careful implementation of a consensus protocol that relieves the
CPU of the responsibility of data movement permits a high performance
implementation entirely in software that addresses the order of magnitude
network related performance gap in Figure~\ref{fig:problem} using only commodity
10 Gigabit ethernet.

The rest of the paper is organized as follows. We describe Cyclone's system
architecture in Section~\ref{sec:sysarch}. We describe this two level log
structure in Section~\ref{sec:storage}. The set of optimizations to the network
stack is decribed in Section~\ref{sec:network}. Both the storage (NVM, SSD) as
well as the NIC expose significant parallelism that is left unused when Cyclone
is used to replicate a single log. To exploit this parallelism we use Cyclone to
replicate multiple logs, each using a different instance of the RAFT consensus
protocol. We show how to exploit the commutativity inherent in key-value store
interfaces to make use of this parallelism in Cyclone's design in
Section~\ref{sec:parallelism}.  A detailed evaluation of Cyclone's replication
performance is provided in the context of RocksDB - a popular persistent key
value store - in Section~\ref{sec:evaluation}. We then discuss related work
before concluding.

\section{System Architecture}
\label{sec:sysarch}
A persistent key value store durably stores key-value mappings accessible through a
simple interface:

\begin{itemize}
\item GET(K): returns value corresponding to key K
\item PUT(K, V): sets the value of key K to value V
\end{itemize}

A key value store typically serves multiple client requests concurrently and is
linearizable~\cite{linearizability}. This means any history of requests and
responses can be rearranged such that responses immediately follow corresponding
requests and if a response precedes a request in the original history it also
does so in the rearranged history. Key value stores such as RocksDB also
support atomic writes to multiple keys and snapshots.

Cyclone integrates with key value stores as both a client and server side
library. Cyclone replicates the key value store across a set of replica
servers. The client side library sends \emph{all} requests to a distinguished
leader replica.  On the server side Cyclone accepts requests from clients and calls
into the key value store. It then returns a response to the client. Before
executing any PUT request, Cyclone appends it to a durable log and replicates
the request to the logs of follower replicas. The request is considered
replicated once acknowledged by a majority of replicas. Follower replicas apply
updates from the log in order.

Cyclone uses RAFT~\cite{raft} to keep the logs in sync across the replicas. On a
failure of the leader replica a new leader is automatically elected - RAFT
ensures that the new leader has the most up to date log. The Cyclone client
library automatically locates the new leader on a failover - clients therefore
see only a brief interruption in service. Although read requests (GET
operations) are not logged, they are serviced in FIFO order with the write requests
(PUT operations) currently in Cyclone's log at the moment the read request is
received. This ensures linearizability in the face of failovers.

\section{Storage}
\label{sec:storage}
Storage devices such as flash-based SSDs export a block IO interface. Appending
a small update to the write-ahead log requires a synchronous access to the
SSD. Even with write coalescing using a write cache on the SSD, the round trip
time to a current generation NVMe SSD such as the one we use in this
paper~\cite{ssd_spec}, is of the order of 20 us, limiting the throughput to
under 50K ops/sec. This leads to a high baseline latency at even low load, as
shown in Figure~\ref{fig:problem}. This problem leads many system designers to
underprovision key-value store shards ensuring that it operates at moderate to
high load, using group-commit to batch additions to the log. They therefore pay
a price in latency to ensure the bandwidth to storage is fully utilized.

Cyclone adopts a different strategy by making use of novel memory technologies
to obviate the need to make this tradeoff in the first place. We assume a small
amount of non-volatile memory directly addressable by the CPU rather than being
placed behind a block oriented IO interface. Log appends are done to the ``NVM
log'' placed in this non-volatile memory. While we designed Cyclone to leverage
new non-volatile memory technologies that might be available in the
future~\cite{3dxpoint}, current non-volatile memory options such are NVDIMMs are
more expensive that DRAM. Therefore, for Cyclone to be economical we cannot
assume that a sufficient amount of NVM is available to hold the entire log. The
log entries can only be garbage collected when the in-memory component of the
key value store is checkpointed to secondary storage. The growing size of
available DRAM means that also provisioning sufficient NVDIMMs to hold the
corresponding log of updates is not feasible given that NVDIMMs are
significantly more expensive than standard volatile DIMMs.  To illustrate this,
at the time of writing this paper, an 8 GB DDR4 NVDIMM was retailing at \$612,
while a plain 8 GB DD4 DIMM was retailing at \$178.  We therefore resort to
using a second larger log level placed on a flash SSD. Entries are drained from
the NVM log to this flash log in conveniently large units that are a multiple of
4KB - the optimum IO unit for flash SSDs. This ensures we get the best possible
throughput from a flash device without paying the price of synchronous IO for
small key-value pairs. At the same time, the amount of non-volatile memory
required is small enough to not add undue cost of the server.

The flashlog is written out in segments of configurable size (we use 128KB
segments). A segment buffer is prepared in memory (volatile DRAM) using the
layout shown in Figure~\ref{fig:flashlog_page}. We do not allow objects in the
flashlog to cross a 4KB boundary - linking multiple objects together with a
special flag encoded into the size if necessary. We flush log segments out to
the log file using asynchronous direct IO, and therefore we continue to fill log
segment buffers while keeping IO to previously filled buffers outstanding to the
flash drive. We allocate enough buffers to keep a maximum of 32 outstanding
requests to the SSD. To avoid having to do a synchronous metadata flush, we preallocate (using the
posix\_fallocate call) a gigabyte worth of zero filled disk pages at the end of
the log file whenever we hit its end.

In order to recover from a crash, we make two important assumptions about the
underlying SSD. First, we assume that 4KB is the \emph{minimum} atomic unit for
updating pages on the SSD even under power failure i.e. there are no shorn
writes (otherwise known as torn writes) on a 4KB page~\cite{shorn_writes}. We
also assume that the SSD has power loss data protection meaning that writes
cached in the drive's volatile cache are written to the SSD using a backup
capacitor in the event of power failure - a property of many data center class
SSDs today, including the Intel DC P3600 SSD~\cite{ssd_spec} we use in our
evaluation. Together, these two assumptions mean that we can recover a consistent
prefix of the log on a power failure.

We move log entries from the head of the NVM log to the flashlog buffers in FIFO
order. The NVM log entry is only actually removed when the IO for the
corresponding flashlog page is complete. This means that during recovery we can
have the same log entry both in the NVM log and the flashlog, a condition that
can be detected by examining the log sequence number that we embed in each log
entry.

\begin{figure}
  \centering
  \includegraphics[scale=0.4]{figures2/flashlog_page.pdf}
  \caption{Flashlog segment}
  \label{fig:flashlog_page}
\end{figure}

The two level log design in Cyclone lays the foundation for improvements to the
network stack that we discuss next - a significant contributor to the overheads
in Figure~\ref{fig:problem}. We use RAFT to replicate the NVM log from the
leader replica to the followers. Each replica \emph{independently} drains
committed (in terms of RAFT consensus) entries from the NVM log to the
flashlog. Block IO therefore is no longer a consideration when optimizing data
movement over the network, unlike systems that need to deal with such
problems~\cite{reflex}. Replication reduces to efficiently sending a block of
data already present in directly attached memory over the network.

\section{Network}
\label{sec:network}

The core network operation in replication is to receive a request from the
client at the leader replica and send that exact same request (as a log entry)
to all follower replicas. The leader therefore \emph{multicasts} a received
request packet to follower replicas, a problem well studied in the networking
community when building software packet switches. Software packet switches reach
impressive speeds of millions of packets forwarded per second~\cite{dpdk_perf}, a
number far in excess of the few hundreds of packets we manage in
Figure~\ref{fig:problem}. We therefore proceed by approximating a software
packet switch for the networking component of Cyclone. However software packet
switching does not need to deal with failure - usually requiring only a simple
decision for each packet at the node. RAFT introduces significant
protocol level work for each forwarded packet a challenge that we need to deal
with to approach the desired software packet switching efficiency in the common
case.

\subsection{DPDK}
The Data Plane Development Kit~\cite{dpdk} provides low latency userspace access
to an ethernet NIC, permitting the application to directly send and receive raw
ethernet frames via the transmit and receive queues on the NIC. DPDK is often
used by developers of software packet switches and therefore we also leverage
DPDK as a library for building Cyclone.

DPDK does not by itself provide a TCP stack. This is not a problem. RAFT (and
indeed most consensus protocols) tolerate network losses and reordering by
design due to the need to support asynchronous communication. In addition, most
datacenter networks rarely drop or reorder packets and provide full bisection
bandwith between servers that might serve as Cyclone replicas. We therefore
jettisoned TCP and chose to send raw IP packets encapsulated in ethernet
frames. We currently follow this communication model both for server to server
communication between replicas as well as client server communication for
requests and responses. Although a detailed evaluation is made later, switching
from the kernel TCP/IP stack to DPDK reduces the latency between machines in our
testbed from {\tt 18 us} to as low as {\tt 5 us}, providing a significant boost
to performance.

We direct DPDK to use pages backed by NVM for packet buffers. DPDK uses a
concurrent memory allocator based on reference counting - used by both the NIC
driver and CPU cores. This means that packet allocation and deallocation does
not happen in the same order as their corresponding position in the replicated
RAFT log. To deal with this, we use a level of indirection as shown in
Figure~\ref{fig:nvm_log}. The NVM log is maintained as a circular log of fixed
sized pointers. Adding a level of indirection in the NVM log allows us to
separate the FIFO ordered circular log being manipulated by RAFT from packet
data being managed by the memory allocator of DPDK. Both the circular log and
packet data are in NVM. An advantage of this scheme is that it makes recovery
from NVM easy - appends to the circular pointer log are atomic and we can use
the pointer log to recover allocator state i.e. what pieces of NVM are currently
in use by the log.

\begin{figure}
\centering
\includegraphics[scale=0.3]{figures2/nvm_log.pdf}
\caption{NVM log structure}
\label{fig:nvm_log}
\end{figure}

\subsection{Eliminating Data Movement}
\label{sec:dm}
Software packet switches built on DPDK try to touch as little of the packet data
as possible, minimizing movement of data up and down the cache and memory
hierarchy. Log replication looks like multicast that does not require deep
packet inspection. Software packet switches implementing multicast therefore
simply manipulate packet headers to produce new packets to send on the output
ports. We designed our implementation of raft around the same principle.

\begin{figure}
\begin{verbatim}
    event_receive_client_req()
    {
      if(!check_is_leader()) {
        drop_msg
        return
      }
      Prepend raft log term and index
      Persist to local log
      Transmit to follower replicas
    }
\end{verbatim}
\caption{Cyclone event handling}
\label{fig:control_plane}
\end{figure}

The pseudocode in Figure~\ref{fig:control_plane} describes part of the packet
handling code in Cyclone organized as event handlers triggered on receiving a
packet at the leader. We focus on only one key event for brevity: the event
where a request is received from a client. The first step is to check that this
replica is indeed the RAFT leader in the current view (term). If not, the
message is simply dropped (a timeout causes the client to try a different
server). It then prepends RAFT related information to the packet - this includes
the current term and log index for this entry. Next, it appends a pointer to the
log in Figure~\ref{fig:nvm_log} effectively appending the packet to the
persistent log. Finally, it transmits the packet to follower replicas.

This entire process is done without making any copies of the received packet.
To illustrate how this is achieved, Figure~\ref{fig:packet_layout} shows how
Cyclone manipulates packet layouts across the two steps of prepending a RAFT
header and transmitting to follower replicas. DPDK describes packets using an
``mbuf'' data structure. Roughly speaking, an ``mbuf'' consists of a flat array
of bytes actually containing the packet and a fixed size piece of external
metadata that describes various aspects of the packet, most crucially a pointer
to the start and end of the packet in the byte array. DPDK's userspace drivers
receive packets from the NIC such that they are offset in the byte array by a
configurable amount referred to as ``headroom''. We strip off the existing
network headers in the packet and prepend RAFT related information specific to
each log entry in the headroom by shifting the start pointer appropriately.
These operations are standard enough for software packet switches that DPDK
provides convenient library calls for it. For the final step, we need to prepare
the packet for transmission to the various follower replicas. To do this we
prepare a different packet containing an ethernet header for \emph{each}
targeted replica and ``chain'' the data packet to each of these headers. Each
header is then separately handed off to the driver for transmission via the NIC,
carrying the data packet with it by association.

\begin{figure}
  \centering
  \includegraphics[scale=0.35]{figures2/network_packet.pdf}
  \caption{Cyclone network packet layout}
  \label{fig:packet_layout}
\end{figure}

We now turn our attention to the persistence step in
Figure~\ref{fig:control_plane}. RAFT requires that the log entry be persisted
before it is multicast out to follower replicas. DPDK userspace NIC drivers
operate in DDIO mode~\cite{ddio} where the packet is directly written into the
CPU cache rather than first being DMAed into DRAM and then fetched by the CPU on
demand. Since the NVM is directly attached we need to persist the packet to it
by executing a cacheline flush ({\tt clflush}) instruction for every cacheline
in the packet and the pointer in the pointer buffer to persist these via the
memory bus. This is not too onerous a burden because we can use the newly
introduced {\tt clflush-opt}~\cite{clflush_opt} instruction specifically
intended to efficiently flush to persistent memory without the overhead of the
serialization normally introduced by {\tt clflush}. This allows us to hit full
memory bandwidth on present generation platforms, a quantity in excess of 200
Gb/s per core, well above the near term speeds of network interface cards. We
execute a single serializing {\tt sfence} before the sequence of {\tt clflushes}
to make sure any dirty cacheline due to header manipulation related writes from
the CPU are sent to cache.

\begin{figure}
  \centering
  \includegraphics[scale=0.4]{figures2/batching.pdf}
  \caption{Batching}
  \label{fig:batching}
\end{figure}

Although RAFT is an efficient consensus protocol in the common case, the
protocol state machine still adds significant overhead to each packet, relative
to the time for the packet to flow in from the NIC and back out to the replicas.
We address this problem for the \emph{loaded} case using batching - treating a
whole sequence of client commands as a single RAFT log entry, while avoiding any
copies to group these packets together. Figure~\ref{fig:batching} illustrates
how this is done. We use a burst receive call available in the DPDK userspace
driver to receive a burst of client packets at a time. We then chain these
packets together and treat them as a single log entry from the perspective of
RAFT, amortizing the control plane overheads over the packets (at most 32 at a
time due to current driver limitations). Crucially, batching in Cyclone does not
involve a latency-throughput tradeoff like in many other
systems~\cite{ix-dataplane}. The batch receive call we use in DPDK returns
immediately with whatever number of packets is available, including zero. We
always flush the transmit buffer after every call to DPDK to transmit packets to
replicas. Therefore, we never tradeoff latency for throughput when batching.

Finally, we discuss our choice of RAFT as a consensus protocol. A key advantage
of RAFT is that unlike alternatives such as MultiPaxos~\cite{multipaxos} or
Quorum based replication~\cite{quorum}, the leader is guaranteed to have the
most up to date logs. There is therefore no case where a leader needs to receive
log entries to ``fill'' holes in its logs from follower replicas, simplifying
the protocol state machine we need to implement. A leader can immediately reject
any responses that are not more current than its view (term). In turn, the
persistent log operates as a double ended queue, with entries either being
appended (or possibly deleted at followers) at the end and being deleted from
the front (after commit). This is critical to efficient operation of Cyclone, as
it allows a simple top level structure (a circular buffer of pointers) to
represent the log.

\section{Parallelism}
\label{sec:parallelism}

%\begin{figure}
%  \centering
%  \includegraphics[scale=0.35]{figures2/race.pdf}
%  \caption{Race with multiple physical logs}
%  \label{fig:race}
%\end{figure}

Although the work described in the previous section significantly boosts the
network performance of replication in Cyclone, it still does not come close to
saturating the capabilities of even the commodity network interface card that we
use. This is because NICs today encapsulate significant parallelism in terms of
multiple send and receive queue pairs. Exploiting this capability to improve
throughput requires us to remove the bottleneck of a single sequential log in
RAFT. To this end, we extend the implementation described thus far to run
multiple copies of the RAFT consensus protocol each maintaining and replicating
its own two level log. We now refer to these logs as physical logs. All these
instances however exist in the same process address space as the key value store
application itself and therefore manipulate the same application.

The first key question we deal with is - how do Cyclone clients decide which
physical log to send a {\tt PUT} or {\tt GET} request to ? A key constraint here
is that all replicas should agree on the seqeunce of updates to a key - as
otherwise they would diverge in terms of state. Allowing updates to the same key
to go to different physical logs would lead to a divergence as there is no
guarantee about the order in which the independent RAFT instances would apply
the update. Therefore all updates to a key must go to the same physcial
log. A similar constraint applies to reads in order to ensure that any committed
writes are seen by readers. Allowing reads to a key to go to a different RAFT
instance than the write means that it could be served on a machine where the
write is yet to be applied but where the machine is the leader for the RAFT
instance. All reads and write to a key must therefore go to the same physical
log. We achieve this by hashing the key to select the physical log.

It turns out that this simple \emph{necessary} scheme is also \emph{sufficient}
to ensure linearizability in the face of failure. Clearly the system is
linearizable for a single key since it involves only a single physical log. In
addition, one can switch operations to different keys in a history without
affecting the correctness of the history because operations to different keys
are commutative. Any history of the system is therefore linearizable.

Multiple physical logs in Cyclone operate in a shared nothing manner by
partitioning the NVM and SSD space evenly between them and by allocating
dedicated NIC queue pairs to each instance of RAFT. This works well because the
memory hierarchy including the NVM and the SSD efficiently support concurrent
operations. The only synchronization necessary is when doing reads or writes to
the single shared key value stores.  The level of concurrency therefore is
constrained only by the concurrency available in the software architecture of
the persistent key value store itself, for which a number of good designs
already exist~\cite{flodb}.

In our discussion thus far, we have focused on operations to a single
key. Supporting atomic writes or reads to multiple keys while maintaining
linearizability requires special handling with multiple physical logs so as not
to break linearizability - that we discuss next.

\subsection{Ganged Operations}
Supporting reads or writes to multiple keys that map to different physical logs
requires a ganged operation across those logs such that all the operations
complete together in the face of failures. Since our starting point is a single
key value store, we constrained Cyclone's design early on to avoid the need for
any two phase commit.

The first key problem we need to deal with is that the leaders for the different
RAFT instances can be located on different machines. To avoid having a client
co-ordinate different machines for a single operation, we constrain the RAFT
leader election algorithm for the different RAFT instances to converge to the
same machine. We do this by triggering re-elections for a RAFT instance other
than that of the first physical log until it is on the same machine as the
\emph{first} physical log. This process is therefore resilient to failed
machines. RAFT requires that the leader have the most up to date log. To ensure
that the process converges we therefore ensure that the currently elected leader
brings a majority quroum up to date before triggering a relection. We also do
not accept any requests from clients until all leaders are co-located. This
process is guaranteed to converge as long as all no network partitions occur
\emph{within a machine} i.e. some RAFT instances are able to send and receive
packets, while others are not, a possiblity we discount due to our single
process design.

Next, we need to simultaneously inject the ganged operation into all
participating physical logs on the machine hosting the leaders. Clients always
dispatch batched writes to a fixed co-ordinating physical log/RAFT instance
(henceforth called the co-ordinator), which is then responsible for forwarding
the request to the participating logs. We make use of packet cloning primitives
to avoid making a physical copy of the packet, generating indirect references
instead. The co-ordinator also adds a ``nonce'' - a unique timestamp to the
packet. In addition, the co-ordinator adds a unique view number to the packet
containing the term numbers of all the participating RAFT instances (read from
shared memory of the co-located leaders).  The event that applies a ganged
operation is described in the pseudocode of Figure~\ref{fig:ganged_ops}. We
defer the discussion of how we generate the nonce to later in this section. We
also assume that a unique barrier is allocated in shared memory for each ganged
operation. We discuss later how this is done without using dynamic allocation.

The complexity in Figure~\ref{fig:ganged_ops} arises from the need to handle
both shared memory concurrency and failure during replication. Without failure,
Figure~\ref{fig:ganged_ops} is straightforward. A barrier is executed on all
RAFT instances. Once replication is complete on all participating physical logs
(or this is a read ganged operation), a state indicated by all necessary bits
being set in the barrier mask, the RAFT instances simultaneously execute the
ganged operation.

Each RAFT instance always publishes the current term it is operating in. A
failure causes a new leader to be elected moving forward the term. If the
participants of a ganged operatation do not detect this case, they could be left
waiting for ever as ganged operation might never successfully replicate in the
failed over RAFT instance. We detect failure on any RAFT instance by having each
instance publish its current term and continuously comparing the view in the
ganged operation to it. If any RAFT instance has moved past that term, the
ganged operation is then terminated by having one of the still alive participants set a
failure flag and write the appropriate bit in the barrier mask on behalf of the
failed participant, moving execution past the barrier wait - a technique we
borrow from lock-free data structures~\cite{lock-free}.

\begin{figure*}
  \centering
\begin{verbatim}
    // Apply ganged operation
    event_apply_ganged_op (packet, barrier)
    {
      if(co-ordinator)
        atomic set bit me in barrier.mask         
        do
         for each participant p in operation
          if public_data[p].view > packet.view
           barrier.failed = true 
           atomic set bit for p in barrier.mask
        while barrrer.mask != mask of all participants
        if barrier.failed
         Ganged replication failed. Send retry to client. return.
        else       
         execute operation
         send response
      else
        wait until 
         public_data[co-ordinator].view > packet.view OR
         barrier.mask == mask of participating KV cores
        if public_data[co-ordinator].view > packet.view OR
           barrier.failed
         Ganged replication failed. Send retry to client. return.
        else
         execute operation
    }
\end{verbatim}
\caption{Ganged Operation}
\label{fig:ganged_ops}
\end{figure*}

The assumption in Figure~\ref{fig:ganged_ops} is that each ganged operation is
mapped to a unique barrier. We achieve this by using a fixed piece of memory
owned by the co-ordinator to hold the barrier and write the nonce to it in order
to indicate that the barrer is active for the corresponding ganged
operation. Participants watch for the nonce to know when to execute the
ganged operation barrier in Figure~\ref{fig:ganged_ops}, while also monitoring
the leader's published view to detect the case where the ganged operation fails
to replicate on the co-ordinator's physical log.

Finally, we describe how we generate the nonce. The nonce is generated on the
co-ordinator RAFT instance by concatenating the ethernet MAC ID of the first NIC
on the system with a 64 bit value that is the number of CPU timestamp counter
cycles since epoch time (read from the real time clock at startup plus the
number of cycles from the CPU {\tt rdtsc} instruction). The nonce can only be
repeated if the same machine manages to fail and come back up in less time than
the real time clock drift (controlled with NTP), a possibility that we discount.

Ganged operations possibly consititute the most complex part of Cyclone but the
code weighs in at well under a couple of hundred lines. We believe that
this additional complexity is still small compared to distributed transactions
using two-phase commit - especially when taking into account failure recovery -
as used in systems such as FARM~\cite{farm}.

\section{Evaluation}
\label{sec:evaluation}
We evaluate Cyclone on a 12 node x86 Xeon cluster connected via a 10 GigE
switch. Three of the machines are equipped with 1.6TB Intel DC P3600 SSDs and
4*10 GigE ports. The remaining nine machines do not have SSDs and have only one
10 GigE port, serving as clients for most of the experiments. We turn on jumbo
frame support in the 10 GigE switch to enable maximum use of batching in
Cyclone. As with other work~\cite{faast}, we use DRAM on the machines to proxy
for NVDIMMs where necessary - the persistent memory needed never exceeds 64 MB
regardless of the size of the key value store or second level log on flash. We
divide the evaluation into three parts.

First, we evaluate Cyclone's performance with a single level log as a pure
software packet switch. For this purpose, Cyclone uses a dummy server stub that
simply echoes the client request back to it. Next, we evaluate
performance with the dummy stub, but with the addition of the second level of log
on flash. Finally, we evaluate performance when integrated with
Rocksdb~\cite{rocksdb} as an alternative to Rocksdb's write ahead log.

Unless otherwise mentioned, we use a 60 byte header followed by an optional
payload for experiments. We log both the header and payload. Also, unless
otherwise mentioned we use 8 physical logs (and associated RAFT instances) each
mapped to a dedicated core. The remaining 32 cores are dedicated to the stub
server or Rocksdb, as the case may be.

We begin by systematically evaluating network stack related optimizations
applied in Cyclone to replicate the NVM log in Figure~\ref{fig:network_opts} -
with no payload. The y-axis reports latency seen at the client (which means two
network round trips with replication). Using TCP/IP to replicate a RAFT log tops
out at around 30K entries/s.  Switching to DPDK (the line marked +DPDK) improves
the throughput by an order of magnitude to around 500K entries/s. Using batching
(the line marked +batching) improves the performance further bringing us close
to a million entries/s. Scaling to 8 physical logs (+8 phy logs) improves
performance to close to 2M entries/s. Finally using all 4 ethernet ports on the
machine to replicate entries improves performance considerably to 6M
entries/s. In all, performance improves by 200X over the TCP/IP single log
baseline. Cyclone also considerably improves the latency for replication, from
close to 100us with TCP/IP to around 30us at peak throughput.

One can draw two important conclusions from
Figure~\ref{fig:network_opts}. First, one can indeed treat log replication as a
software packet switching problem \emph{provided} the persistent log can be held
in directly attached memory eliminating the block interface from the critical
path. Second, relieving the CPU from the overhead of data copies has a
significant positive effect on performance. Finally, using multiple physical
logs is essential to exploiting the concurrency available at the level of the
network interface card (even a single one!) that would otherwise go wasted due
to the serializing abstraction of a single log. Finally, we note that DPDKs raw
packet forwarding capability on that configuration is close to 12M packets per
second. We do not reach that mark due to the overhead of running the replication
protocol rather than a simple switching decision per packet. Although batching
goes some way in amortizing this overhead it cannot be eliminated entirely
because consensus protocols are more complex than a simple packet switching
decision.

\begin{figure}
\includegraphics[scale=0.6]{results2/network_opts.pdf}
\caption{Network optimizations for top level log}
\label{fig:network_opts}
\end{figure}

Both the replica count and payload size can have significant impact on Cyclone's
network performance. First, the number of replicas dictates the outgoing message
rate from the leader replica and therefore increasing the replication factor can
decrease Cyclone's performance. Figure~\ref{fig:replicas} shows the impact of
varying replica count. Using only a single replica cuts out a network round trip
and shows the best unloaded latency (10 us) and peak throughput (near 10M
entries/s). Adding replicas decreases the peak throughput down to around 2M
entries/s with 5 replicas. We note that a number of previous pieces of
work~\cite{faast, farm} use three replicas and therefore we focus on three
replicas for the replicated cases we consider below. The second factor that
dictates Cyclone's performance is the size of the log entry being
replicated. Figure~\ref{fig:payload} shows the effect of increasing the payload
size from zero to 512 bytes. Peak throughput drops from 6M entries/s to
approximately 2M entries/s. At this replication rate, the leader replica needs
to transmit data at approximately 30 Gbit/s. Coupled with the cost of network
headers all four 10 GigE ports are now saturated and therefore Cyclone hits the
network line rate bottleneck at this point. One can contrast this with the
previous experiment where the bottleneck to throughput was the cost of running
the consensus protocol for batches of packets.

\begin{figure}
\includegraphics[scale=0.6]{results2/replicas.pdf}
\caption{Impact of replica count}
\label{fig:replicas}
\end{figure}

\begin{figure}
\includegraphics[scale=0.6]{results2/512.pdf}
\caption{Impact of payload size}
\label{fig:payload}
\end{figure}

We now turn our attention from the network component of Cyclone to the storage
one by adding the second level flashlog. We evaluate the impact of adding the
flashlog in Figures~\ref{fig:flashlog} and \ref{fig:flashlog_512}. The benefit
of the second level flashlog is that it lets us limit the amount of needed
NVDIMM memory. The hypothesis was that batching the movement of data from the
NVM log to the flashlog would be sufficient to hide the latency of the block IO
device from network speed replication. The results confirm that in terms of
latency for the low to moderate load cases, our two level log arrangement is
effective at hiding the latency of the NVMe SSD. The picture however is
different for peak throughput. For small updates there is no impact on
throughput when we add the SSD. On the other hand 
Figure~\ref{fig:flashlog_512} shows that using a 512 byte payload has a
significant impact on peak throughput - it drops to approximately 350K
ops/sec. This corresponds to around 50K 4KB IOPS to the SSD to write out the
flashlog pages. This is in fact the expected IOPS limit with the 32 outstanding
requests to the drive that we maintain (Section~\ref{sec:storage}).
A final point about Figure~\ref{fig:flashlog_512} is that once we are past the
storage bottleneck the latency spike is dramatic and large enough to trigger
Cyclone's failure detector and repeated retries from the clients. There are -
therefore - no points on the ``knee'' of the curve as in the pure packet
switched one-level log case.

One can therefore conclude that using a small amount of directly attached NVM is
effective at hiding the latency of a background block storage device from the
criticial path of replication. Further, for small updates the throughput
bottleneck is the capability of the software to run the consensus protocol but
shifts to the secondary storage device for larger request sizes.

\begin{figure}
\includegraphics[scale=0.6]{results2/flashlog.pdf}
\caption{Impact of adding second level log}
\label{fig:flashlog}
\end{figure}

\begin{figure}
\includegraphics[scale=0.6]{results2/flashlog_512.pdf}
\caption{Second level log with 512 bytes payload}
\label{fig:flashlog_512}
\end{figure}

We now evaluate ganged operations. The primary purpose of
ganged operations is to avoid the need for distributed transactions to
manipulate what is a single shared memory image and therefore we were most
concerned about unloaded latency given the complexity of synchronizing different
replication quorums as well executing our rendezvous protocol on multiple
cores. We therefore setup an experiment where a single client - reflecting the
unloaded case - made ganged requests to the replicas. We varied the number of
physical log participating in the request from one to the full complement of 8.
Figure~\ref{fig:ganged} shows the results. Unloaded
latency increases slowly as we increase the number of active physical logs - to
around 32 us from the baseline of 21 us. The reason for this increase is that
the DPDK userspace NIC driver pipelines request processing for communication
between the CPU core and the NIC. Simultaneous replicaton on multiple physical
logs represents a worst case for this arrangement as all eight cores try to
send replication messages to their replicas, adding about a microsecond of
serial latency each, to dispatch requests to the NIC. Regardless, the experiment
underlines the value of our ganged operation design over a distributed
transaction that would need 4 network hops for the two phase commit (and
replication) using an external co-ordinator, a minimum of 40us.

\begin{figure}
\includegraphics[scale=0.6]{results2/multi.pdf}
\caption{Ganged Operations}
\label{fig:ganged}
\end{figure}

We now evaluate Cyclone integrated with the Rocksdb persistent key value
store. Rocksdb is a complex industrial strength persistent key value store and
an important target for practical use of Cyclone.  Rocksdb is accompanied by a
complex array of performance tuning knobs to get the best performance from
flash. Designing LSM tree based key value stores to extract good performance
from flash and make best use of DRAM on the machine is an area of active
research~\cite{flodb, triad}. We wished to make sure that our conclusions would
apply even as Rocksdb performance improves with the integration of new
ideas. Keeping this in mind, we configured RocksDB to place all files for the
key value store (SSTables) on a RAMdisk, which presumably represents the limit
in performance for both software enhancements as well as improvements to
flash-based SSDs. However - both Rocksdb's own write ahead log and the
alternative of Cyclone's second level flashlog - are placed on the SSDs. In
effect we therefore make availability a harder problem in this setting,
guaranteeing our design can handle the immediate future.

We use 8 bytes keys with either 8 bytes values or 256 byte values.
Since our focus is replication and only update requests are replicated, we use a
pure update workload to test the capability of the system.

Before evaluating with Rocksdb, we measure the baseline performance of
replicating the log with Cyclone for the given request sizes - Rocksdb performs
a no-op. We note that in addition to the key and value, we are also logging
RocksDB specific request data such as operation type and the request header,
making this different from the previous experiments.
Figure~\ref{fig:kv_baseline} shows the baseline performance for the chosen
request sizes. With the smaller request size, Cyclone can conservatively sustain
close to a million requests a second at a latency of just under 25us. With the
larger request size, Cyclone can sustain around 350K requests a second, again at
a latency of just under 25 us. Armed with these baseline numbers we now examine
how well Cyclone performs with Rocksdb.

\begin{figure}
\includegraphics[scale=0.6]{results2/kv_baseline.pdf}
\caption{Baseline KV replication performance}
\label{fig:kv_baseline}
\end{figure}

The performance of Rocksdb with Cyclone for the small update workload is shown
in Figure~\ref{fig:rocksdb} - essentially presenting the solution to the problem
description graph in Figure~\ref{fig:problem}.  We consider four different
setups. The line labeled 'rocksdb' is the key value store running with no
logging whatsover - a system crash would lead to data loss. The line labeled
'rocksdb/wal' is for Rocksdb running with its write ahead logging turned on. The
large gap between these two is the overhead of the existing Rocksdb WAL
solution. The line labeled 'rocksdb/Cyclone 1 way' is a two level Cyclone log
but without any replication. The line almost exactly tracks the performance of
Rockdb. As suggested by the baseline replication performance, Cyclone is able to
provide a write ahead log with no overhead to Rocksdb. The line labeled
'rocksdb/Cyclone 3 way' is with 3-way replication turned on. Other than a 20us
delta due to the extra network round trip, the line almost exactly tracks
Rocksdb performance with no logging. Cyclone therefore provides high
availability to Rocksdb at a fraction of the cost of its existing single machine
write ahead log. We also repeat the experiment for the larger update size in
Figure~\ref{fig:rocksdb_256}. The conclusions are identical: Cyclone solves
Rocksdb's write ahead logging problem.

\begin{figure}
\includegraphics[scale=0.6]{results2/rocksdb.pdf}
\caption{Rocksdb - small updates}
\label{fig:rocksdb}
\end{figure}

\begin{figure}
\includegraphics[scale=0.6]{results2/rocksdb_256.pdf}
\caption{Rocksdb - large updates}
\label{fig:rocksdb_256}
\end{figure}

Next, we consider the problem of supporting Rocksdb's write-batch operation that
atomically writes a set of key-value pairs into the KV store. This is a
practical application of ganged operations in Cyclone. In order to
perform the operation with Cyclone managing the log, the client must issue a
ganged operation across physical logs owning the keys in the write batch.
A key concern here was whether Cyclone would add any latency to the
operation due to the extra synchronization needed across replication quorums (
shown in a previous experiment).
We examine the problem for the unloaded case
and small updates in Figure~\ref{fig:rocksdb_multi} for a single client with
increasing number of keys in the batch - till 32 keys.
The line labeled Rocksdb is with no write ahead logging. We
note an increasing latency for this baseline indicating Rocksdb itself takes longer
with larger key batches. The existing option of Rocksdb/wal has considerably
larger latency. Cyclone does an effective job of cutting down on this latency
even as it needs to pay a price for synchronizing multiple physical
logs. Cyclone therefore provides effective replicated logging for batched
operations using its idea of ganged operations, while scaling the performance of
single operations using its multiple physical logs.

\begin{figure}
\includegraphics[scale=0.6]{results2/rocksdb_multi.pdf}
\caption{Rocksdb - batched writes}
\label{fig:rocksdb_multi}
\end{figure}

Our evaluation has, thus far, focused on synthetic write heavy
workloads. To illustrate that Cyclone can also be useful for read-heavy
workloads, we configured our clients to generate key value requests that mimic
the distribution of value sizes in Facebook's ETCD
trace~\cite{fb_workload_analysis,fb_scaling_memcache} and are read
heavy (95\% reads). Figure~\ref{fig:fb} illustrates that even with a
read-dominated workload, Cyclone is effective at providing a write ahead log
with 3 way replication with a lower performance overhead than Rocksdb's own
write ahead logging.

\begin{figure}
\includegraphics[scale=0.6]{results2/fb.pdf}
\caption{Rocksdb - read heavy workload}
\label{fig:fb}
\end{figure}

Finally, we showcase the benefit of using Cyclone beyond pure performance as
compared to the existing single machine Rocksdb write ahead log. Cyclone brings
multi-machine availability with the ability to automatically failover. We
demonstrate this in Figure ~\ref{fig:timeline} that shows the timeline of a run
where we kill the server process on the leader replica. Cyclone is configured to
with a 30ms failure detection timeout after which the client library tries
replicas in turn to locate the leader - in this case it fails over in about
60ms.

\begin{figure}
\includegraphics[scale=0.6]{results2/failover.pdf}
\caption{Rocksdb - failover}
\label{fig:timeline}
\end{figure}

\section{Related Work}

Cyclone takes a software only approach to improving replication performance
using commodity network hardware. In contrast, Mojim~\cite{mojim} pairs directly
attached persistent memory with RDMA to replicate persistent data structures. A
comparison is hard since Mojim is designed for the hypothetical scenario where
there is enough persistent memory to hold all persistent data, while Cyclone is
designed for the practical case where both the write-ahead log and key value
store data structures need to be maintained largely on flash. Nevertheless we
can compare their replication performance for the case where we use only the NVM
log in Cyclone. On 40Gbps infiniband, Mojim reported being able to mirror data
at 4GB/s. In contrast, Cyclone can replicate data three ways using 4*10 Gbps
commodity ethernet NICs at a peak of roughly 3.7 GB/s. This is with
Cyclone is also running the RAFT consensus protocol, while Mojim is doing simple
mirroring of main memory. One can conclude that network offload via RDMA and
Infiniband is not an a-priori requirement for high throughput replication. This
is made possible by Cyclone's approach that relieves the CPU from the burden of
data copies. On the other hand, the latency of a hop on commodity ethernet at 5 us with
our DPDK stack is larger than the close to 2us latency with
Infiniband. Depending on the use case for the key value store application the
extra 6us with ethernet for the round trip to the replica might be a concern for
some applications. We note that mapping the commutativity of key value store
interfaces to multiple logs in Cyclone is a technique that can also be applied
to RDMA. Mojim confirms that concurrency is a significant
determinant of performance when replicating over RDMA.

Consensus in a box~\cite{consensus_box} implements the Zookeeper atomic
broadcast protocol together with a NIC on an FPGA. That work ignored durability,
focusing purely on network performance for replication. Latency is
excellent due to the FPGA that cuts out the path from CPU caches to the NIC and
are reported to be as low as 3us for a round trip from leader to follower
replica compared to the 6us we observe with DPDK. Notably however they reported
7us when using TCP/IP from the FPGA rather than their own connection oriented
network protocol. On the other hand their peak replication throughput is 6M
replications/sec on 66 Gbps of aggregate network bandwith, comparable to the
same mark obtained by Cyclone on an aggregate 40 Gbps of network bandwidth when
not using the log on the SSD. We believe the general applicability afforded by
our software only solution together with the fact that we provide persistence by
design makes it a compelling alternative to an FPGA based solution even given
the larger latency.

DARE~\cite{dare} implements state machine replication over RDMA, using a custom
replication protocol. DARE does not consider persistence in its design. It is
worthwhile to point out that since DARE allows any replica to become the leader
it requires a log adjustment protocol for the leader to bring its logs up to
date, a situation we avoid by using RAFT as described in
Section~\ref{sec:dm}. Using 40 Gb/s speeds, and 3-way replication DARE reported
being able to replicate 500K requests/sec. This is without persisting the log of
updates. In contrast, Cyclone replicates upwards of 6 million similarly sized
requests each second with 4*10 Gbps ethernet (without the second level
flashlog). DARE reports an unloaded write latency of 15us, which is lower than
the unloaded write latency of Cyclone at 20us, thanks to infiniband. The
conclusions we can draw are similar to the comparison with Mojim: a carefully
implemented log replication system obviates the need for network offload when
doing replication.

{CRANE}~\cite{crane} is a system for replicating mutithreaded programs using
Paxos. It uses deterministic multi-threading to ensure replicas converge to the
same state as opposed to semantic equivalence in Cyclone using the commutativity
of the key value store interface. Although CRANE is more generally applicable in
that it makes no assumptions about the application API, deterministic
multithreading comes with high overheads for applications that store large
amounts of data - CRANE reports around a 2X slowdown for MySQL, likely rendering
it unsuitable as a replacement for the write ahead log in key value stores.

The scalable commutativity rule~\cite{scalable_commutativity} generalizes the
idea of commutativity in key value store interfaces that we have used for
horizontal scaling in Cyclone. Conveniently, it provides a tool for checking when
API calls can commute in a history. It is a simple exercise to express a key
value store API in COMMUTER and check that operations to different keys
commute. This provides a theoretical basis for horizontal scaling, something we
have omitted for brevity from this paper. The scalable commutativity rule also
suggests that scalable key value store implementations such as
FloDB~\cite{flodb} would necessarily have commutative interfaces, enabling more
general applicability of Cyclone.

There is an in progress transformation of durable memory in datacenters
with the advents of 3DXPoint. Researchers have speculated on the impact of cheap
and high volume directly attached 3DXPoint~\cite{tiering}. Memory vendors have
also begun shipping software libraries such as NVML~\cite{nvml} for building durable
in-memory data structures using undo logging. A durable key value store using
3DXPoint would be an idea target for Cyclone because each NVM log entry can be
discarded as soon as it is applied to the persistent key value store removing
the need for the flashlog. This would allow Cyclone to operate at pure packet
switching speeds, a goal we hope to achieve in the future with a single level
log for NVML data store applications.

Cyclone provides a client library that automatically times out and retries
requests to deal with failure. Protection against repeated execution of the same
request is currently left to the key value store and client, since we felt that
exactly once semantics is optional when updates such as put operations are
idempotent. We have experimented with exactly once semantics as a basis for
distributed transactions across shards running Cyclone integrated key value
stores. This is along the lines of similar work~\cite{raft_lin}, using a fixed
client space and maintaining information about last seen sequence numbers from
clients in NVM. Distributed transactions on Cyclone shards however are outside
the scope of this paper.

\section{Conclusion}
Cyclone addresses two fundamental difficulties in adding high availability to
persistent key value stores. It leverage a small amount of non-volatile memory
to avoid paying the cost of the block IO interface for every update
operation. It also leverages software packet switching principles to replicate
the write ahead log thus avoiding any need for offload capable network
interfaces.  Cyclone and its Rocksdb bindings are slated for release to the open
source community and we hope both the code and the ideas in the system itself
will help key value store builders and researchers deal with high availabilty
without compromising performance.
\newcommand\myurl[2]{\url{#1}}
\bibliographystyle{plain}
\bibliography{paper}

\end{document}